Inferring Graphics Programs from Images

### 2017-05-09 paper.tex
some revision based on Dan's comment

\section{Introducing visual programs}

\textbf{Comment for reader: this paragraph is a little grandiose and goes
  beyond what we've actually done. But these are the kinds of things
  that motivate the work, and I think that in some form these ideas
  should be in the introduction or the conclusion.} How could an agent go from noisy,
high-dimensional perceptual input to a symbolic, abstract object, like
a computer program?  Here we consider this problem within the context
of \emph{graphics program synthesis}.  We develop an approach for
converting natural images, such as hand drawings, into executable
source code for drawing the original image.
\remark{The use of `graphics programs / visual programs' in the paper title, title of this section, and the body of this section feels too broad. `Graphics program' could conjur a lot of different ideas (esp. 3D graphics); don't want to set the reader up to expect one thing and then be disappointed that what you've done isn't that. You bring up diagram-drawing later in the intro; I think it should be made clear sooner (and certainly mentioned explicitly in the abstract, when you get around to writing that).}

\newcommand{\exampleImageSize}{2cm}
\begin{figure}[h]
  \begin{minipage}[t]{0.8\linewidth}  
\begin{tabular}{llll}
  \includegraphics[width = \exampleImageSize]{figures/expert-60.png}&
  \includegraphics[width = \exampleImageSize]{figures/expert-5.png}&
    \includegraphics[width = \exampleImageSize]{figures/expert-17.png}&
    \includegraphics[width = \exampleImageSize]{figures/expert-58.png}\\
  \includegraphics[width = \exampleImageSize]{figures/60.png}&
  \includegraphics[width = \exampleImageSize]{figures/5.png}&
    \includegraphics[width = \exampleImageSize]{figures/17.png}&
    \includegraphics[width = \exampleImageSize]{figures/58.png}
\end{tabular}
\subcaption{}
  \end{minipage}%
  \begin{minipage}[t]{0.2\linewidth}
    \begin{tikzpicture}
      \node(picture) at (0,0) {\includegraphics[width = 2cm]{figures/expert-31.png}};
      \draw[very thick,->] (picture.south)  -- (0,-2);
    \end{tikzpicture}
    \texttt{FOR} $0\leq i\leq 3$:\\
    \texttt{RECTANGLE} $()$\\
    \texttt{FOR} $0\leq j < i$:\\
    \texttt{CIRCLE} $()$\\
    \subcaption{}
  \end{minipage}
  \caption{(a): Model parses hand drawings (top) into \LaTeX~(bottom). (b) Synthesizes high-level \emph{graphics program} from hand drawing.}
  \end{figure}

High dimensional perceptual input is ill matched to the abstract
semantics of a programming language. But programs with constructs like
recursion or iteration produce a simpler \emph{execution trace} of
primitive actions. Our hypothesis is that the execution trace of the
program is better aligned with the perceptual input, and that the
trace can act as a kind of bridge between perception and programs. We
test this hypothesis by developing a model that learns to map from an
image to the execution trace of the graphics program that drew it.
With the execution trace in hand, we can bring to bear techniques from
the program synthesis community to recover the latent graphics program.
\remark{This is an \emph{excellent} explanation! I would add maybe one more sentence / citation to elaborate on what you mean by `techniques from the program synthesis community', as this will be an unfamiliar concept to some readers, I imagine.}

In this work we consider programs that draw diagrams, similar to those found in papers.

@@

Figure~\ref{architecture} illustrates this architecture.  We first
pass the target image and a rendering of the trace so far to a
convolutional network. Given the features extracted by the
convolutionional network, a multilayer perceptron then predicts a
distribution over the next drawing command to add to the trace.  We
predict the drawing command token-by-token, and condition each token
both on the image features and on the previously generated tokens.
For example, the network first decides to emit the \verb|CIRCLE| token
conditioned on the image features, then it emits the $x$ coordinate of
the circle conditioned on the image features and the \verb|CIRCLE|
token, and finally it predicts the $y$ coordinate of the circle
conditioned on the image features, the \verb|CIRCLE| token, and the
$x$ coordinate.
\remark{There are some more details that are important to provide about this architecture, though possibly in an Appendix: the functional form(s) of the probability distributions over tokens, the network layer sizes, which MLPs share parameters, etc.}

@@

We train the network by sampling execution traces $T$ and target
images $I$ for randomly generated scenes, and maximizing
(\ref{objective}) wrt $\theta$ by gradient ascent. Despite the
architecture being recurrent, training is fully supervised. In a
sense, this model is like an autoregressive variant of AIR.
\remark{I like that you make this connection, but it could be made more precisely. Specifically, (1) the architecture isn't \emph{really} recurrent (it uses no hidden state cells), so it'd be good to use a different term or drop this part of the point: (2) training of recurrent nets is also typically fully-supervised (Most RNNs lack latent variables per timestep)---if you're thinking about AIR specifically, maybe just say that, and (3) it's like an autogressive AIR \emph{without attention}.}
\remark{Something related to this that's also cool to point out: training this model doesn't require backpropagation across the entire sequence of drawing commands (drawing to the canvas `blocks' the gradients, effectively offloading memory to an external (visual) store, so in principle it might be scalable to much longer sequences.}

@@

  \node[draw,thick,circle,minimum size = 1cm](c1) at ([xshift = 1.5cm]features.north);
  \node(l1) at ([yshift = -2cm]c1.south) {\verb|CIRCLE(|};
  \node[draw,thick,circle,minimum size = 1cm](c2) at ([xshift = 1cm]c1.east);
  \draw[->,thick] (c1.south) -- (l1.north);
  \node(l2) at ([yshift = -2cm]c2.south) {\verb|X=3,|};
  \draw[->,thick] (c2.south) -- (l2.north);
  \node[draw,thick,circle,minimum size = 1cm](c3) at ([xshift = 1cm]c2.east);
  \node(l3) at ([yshift = -2cm]c3.south) {\verb|Y=14)|};
  \draw[->,thick] (c3.south) -- (l3.north);

@@

\begin{table}
\begin{tabular}{ll}\toprule
  \begin{tabular}{l}
    \verb|CIRCLE|$(x,y)$
    \end{tabular}& Circle at $(x,y)$\\
  \begin{tabular}{l}
    \verb|RECTANGLE|$(x_1,y_1,x_2,y_2)$
    \end{tabular}&Rectangle with corners at $(x_1,y_1)$ \& $(x_2,y_2)$\\
  \begin{tabular}{l}
    \verb|LINE|$(x_1,y_1,x_2,y_2,$\\
    \hspace{1cm}$\text{arrow}\in\{0,1\},\text{dashed}\in\{0,1\})$
    \end{tabular}&Line from $(x_1,y_1)$ to  $(x_2,y_2)$, optionally with an arrow and/or dashed\\
  \begin{tabular}{l}
    \verb|STOP|
  \end{tabular}&Finishes execution trace inference
\\  \bottomrule
\end{tabular}
\caption{The deep network in (\ref{architecture}) predicts drawing commands, shown above.}
\label{drawingCommandTable}
\end{table}

@@

\begin{table}
  \begin{tabular}{rl}
  Program$\to$&Command; $\cdots$; Command\\
  Command$\to$&\texttt{CIRCLE}(Expression,Expression)\\
  Command$\to$&\texttt{RECTANGLE}(Expression,Expression,Expression,Expression)\\
  Command$\to$&\texttt{LINE}(Expression,Expression,Expression,Expression,Boolean,Boolean)\\
  Command$\to$&\texttt{FOR}$(0\leq \text{Var}  < \text{Expression})$\texttt{ \{ }Program\texttt{ \}}\\
  Command$\to$&\texttt{REFLECT}$(\text{Axis})$\texttt{ \{ }Program\texttt{ \}}\\
  Expression$\to$&\mathcal{Z}\texttt{ * }Var\texttt{ + }\mathcal{Z}\\
  Var$\to$&A free (unused) variable\\
  \mathcal{Z}$\to$&an integer\\
  Axis$\to$&\texttt{X = }\mathcal{Z}\\
  Axis$\to$&\texttt{Y = }\mathcal{Z}
  \end{tabular}
  \caption{Grammar over graphics programs. We allow loops (\texttt{FOR}), vertical/horizontal reflections (\texttt{REFLECT}), and affine transformations (\mathcal{Z}\texttt{ * }Var\texttt{ + }\mathcal{Z}).}
  \end{table}

>>>>

\section{Introduction}

 How could an agent go from noisy, high-dimensional perceptual input
 to a symbolic, abstract object, like a computer program?  Here we
 consider this problem within a graphics program synthesis domain.  We
 develop an approach for converting natural images, such as hand
 drawings, into executable source code for drawing the original image.
 The graphics programs in our domain draw simple figures like those found in
 machine learning papers (see Fig.\ref{firstPageExamples}).
 \remark{The use of `graphics programs / visual programs' in the paper
   title, title of this section, and the body of this section feels
   too broad. `Graphics program' could conjur a lot of different ideas
   (esp. 3D graphics); don't want to set the reader up to expect one
   thing and then be disappointed that what you've done isn't
   that. You bring up diagram-drawing later in the intro; I think it
   should be made clear sooner (and certainly mentioned explicitly in
   the abstract, when you get around to writing that).}

\newcommand{\exampleImageSize}{2cm}
\begin{figure}[h]
  \begin{minipage}[t]{0.7\linewidth}  
\begin{tabular}{llll}
  \includegraphics[width = \exampleImageSize]{figures/expert-60.png}&
  \includegraphics[width = \exampleImageSize]{figures/expert-5.png}&
    \includegraphics[width = \exampleImageSize]{figures/expert-17.png}&
    \includegraphics[width = \exampleImageSize]{figures/expert-58.png}\\
  \includegraphics[width = \exampleImageSize]{figures/60.png}&
  \includegraphics[width = \exampleImageSize]{figures/5.png}&
    \includegraphics[width = \exampleImageSize]{figures/17.png}&
    \includegraphics[width = \exampleImageSize]{figures/58.png}
\end{tabular}
\subcaption{}
  \end{minipage}%
  \begin{minipage}[t]{0.3\linewidth}
    \begin{tikzpicture}
      \node(picture) at (0,0) {\includegraphics[width = 2cm]{figures/expert-31.png}};
      \draw[very thick,->] (picture.south)  -- (0,-2);
    \end{tikzpicture}\\
    \texttt{for} $0\leq i < 4$:\\
    \texttt{rectangle} $(-3i + 9,2i - 2, -3i + 11,6)$\\
    \texttt{for} $0\leq j < i$:\\
    \texttt{circle} $(3i-2,-2j+5)$\\
    \subcaption{}
  \end{minipage}
  \caption{(a): Model learns to convert hand drawings (top) into \LaTeX~(bottom). (b) Synthesizes high-level \emph{graphics program} from hand drawing.}\label{firstPageExamples}
  \end{figure}

High dimensional perceptual input may seem ill matched to the abstract
semantics of a programming language. But programs with constructs like
recursion or iteration produce a simpler \emph{execution trace} of
primitive actions; for our domain the primitive actions are drawing
commands. Our hypothesis is that the execution trace of the program is
better aligned with the perceptual input, and that the trace can act
as a kind of bridge between perception and programs. We test this
hypothesis by developing a model that learns to map from an image to
the execution trace of the graphics program that drew it.  With the
execution trace in hand, we can bring to bear techniques from the
program synthesis community to recover the latent graphics program.
\remark{This is an \emph{excellent} explanation! I would add maybe one
  more sentence / citation to elaborate on what you mean by
  `techniques from the program synthesis community', as this will be
  an unfamiliar concept to some readers, I imagine.}

@@

Figure~\ref{architecture} illustrates this architecture.  We first
pass the target image and a rendering of the trace so far to a
convolutional network. Given the features extracted by the
convolutionional network, a multilayer perceptron then predicts a
distribution over the next drawing command to add to the trace.  We
predict the drawing command token-by-token, and condition each token
both on the image features and on the previously generated tokens.
For example, the network first decides to emit the \verb|circle| token
conditioned on the image features, then it emits the $x$ coordinate of
the circle conditioned on the image features and the \verb|circle|
token, and finally it predicts the $y$ coordinate of the circle
conditioned on the image features, the \verb|circle| token, and the
$x$ coordinate.
\remark{There are some more details that are important to provide about this architecture, though possibly in an Appendix: the functional form(s) of the probability distributions over tokens, the network layer sizes, which MLPs share parameters, etc.}

@@

We train the network by sampling execution traces $T$ and target
images $I$ for randomly generated scenes, and maximizing
(\ref{objective}) wrt $\theta$ by gradient ascent.
Training does not require backpropagation across the entire sequence of drawing commands:
drawing to the canvas `blocks' the gradients,
effectively offloading memory to an external visual store.
In a sense, this model is like an autoregressive variant of AIR~\cite{eslami1603attend} without attention.

\remark{I like that you make this connection, but it could be made more precisely. Specifically, (1) the architecture isn't \emph{really} recurrent (it uses no hidden state cells), so it'd be good to use a different term or drop this part of the point: (2) training of recurrent nets is also typically fully-supervised (Most RNNs lack latent variables per timestep)---if you're thinking about AIR specifically, maybe just say that, and (3) it's like an autogressive AIR \emph{without attention}.}
\remark{Something related to this that's also cool to point out: training this model doesn't require backpropagation across the entire sequence of drawing commands (drawing to the canvas `blocks' the gradients, effectively offloading memory to an external (visual) store, so in principle it might be scalable to much longer sequences.}

@@

  \node[draw,thick,circle,minimum size = 1cm](c1) at ([xshift = 1.5cm]features.north);
  \node(l1) at ([yshift = -2cm]c1.south) {\verb|circle(|};
  \node[draw,thick,circle,minimum size = 1cm](c2) at ([xshift = 1cm]c1.east);
  \draw[->,thick] (c1.south) -- (l1.north);
  \node(l2) at ([yshift = -2cm]c2.south) {\verb|X=3,|};
  \draw[->,thick] (c2.south) -- (l2.north);
  \node[draw,thick,circle,minimum size = 1cm](c3) at ([xshift = 1cm]c2.east);
  \node(l3) at ([yshift = -2cm]c3.south) {\verb|Y=14)|};
  \draw[->,thick] (c3.south) -- (l3.north);

@@

\begin{table}
\begin{tabular}{ll}\toprule
  \begin{tabular}{l}
    \verb|circle|$(x,y)$
    \end{tabular}& Circle at $(x,y)$\\
  \begin{tabular}{l}
    \verb|rectangle|$(x_1,y_1,x_2,y_2)$
    \end{tabular}&Rectangle with corners at $(x_1,y_1)$ \& $(x_2,y_2)$\\
  \begin{tabular}{l}
    \verb|LINE|$(x_1,y_1,x_2,y_2,$\\
    \hspace{1cm}$\text{arrow}\in\{0,1\},\text{dashed}\in\{0,1\})$
    \end{tabular}&Line from $(x_1,y_1)$ to  $(x_2,y_2)$, optionally with an arrow and/or dashed\\
  \begin{tabular}{l}
    \verb|STOP|
  \end{tabular}&Finishes execution trace inference
\\  \bottomrule
\end{tabular}
\caption{The deep network in (\ref{architecture}) predicts drawing commands, shown above.}
\label{drawingCommandTable}
\end{table}

@@

\begin{table}
  \begin{tabular}{rl}
  Program$\to$&Command; $\cdots$; Command\\
  Command$\to$&\texttt{circle}(Expression,Expression)\\
  Command$\to$&\texttt{rectangle}(Expression,Expression,Expression,Expression)\\
  Command$\to$&\texttt{LINE}(Expression,Expression,Expression,Expression,Boolean,Boolean)\\
  Command$\to$&\texttt{for}$(0\leq \text{Var}  < \text{Expression})$\texttt{ \{ }Program\texttt{ \}}\\
  Command$\to$&\texttt{REFLECT}$(\text{Axis})$\texttt{ \{ }Program\texttt{ \}}\\
  Expression$\to$&\mathcal{Z}\texttt{ * }Var\texttt{ + }\mathcal{Z}\\
  Var$\to$&A free (unused) variable\\
  \mathcal{Z}$\to$&an integer\\
  Axis$\to$&\texttt{X = }\mathcal{Z}\\
  Axis$\to$&\texttt{Y = }\mathcal{Z}
  \end{tabular}
  \caption{Grammar over graphics programs. We allow loops (\texttt{for}), vertical/horizontal reflections (\texttt{REFLECT}), and affine transformations (\mathcal{Z}\texttt{ * }Var\texttt{ + }\mathcal{Z}).}
  \end{table}
